\mythesistitle

### 2011-11-23 src/synopsis.tex
almost done. a wee bit more and i can submit it. almost submitted it today but i have a bus to catch. wooooohooooo bangalore, here i come.

\def \un {\bf 1}

@@

\section{Introduction}
\label{sec:intro}

% \tnote{testing 1 2 3...}
% \tnote[bogus1]{testing 1 2 3...}
% \tnote[bogus2]{testing 1 2 3...}
% \tnote[GTC]{testing 1 2 3...}
% \tnote[bogus4]{testing 1 2 3...}
% \tnote[TH]{testing 1 2 3...}

\subsection{Illustration of the problem}
\label{sec:problem}

% \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
% \def \GGG {{\em ``{\bf G}irl with Curious Hair''}} 
% \def \BBB {{\em ``{\bf B}rief Interviews with Hideous Men''}}
% \def \TTT {{\em ``Everything and More: A Compact His{\bf T}ory of Infinity''}} 
% \def \xLLL {{\bf L}}
% \def \xGGG {{\bf G}} 
% \def \xBBB {{\bf B}}
% \def \xTTT {{\bf T}}

@@

\def \figtabsize {\footnotesize}

@@

% \begin{figure}[t]%[htbp]
%   \centering
%   \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
%   \caption{{\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
%   \label{fig:streetmappath}
% \end{figure}
% \begin{figure}[t] % [h]
%   \centering
%   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
%   \caption{Individual allocation of apartments to students in
%     {\residenceblock} that meets the requirements stated before.}
%   \label{fig:streetmappathpeople}
% \end{figure}
%%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
\begin{figure}[htbp] %[t]%
 \centering
  (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
  (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
 \label{fig:streetmap}
  \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
    allocated. Routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
    \textcolor{green}{$\xTTT$} group}
\end{figure}
\begin{figure}[htbp] %[t]%
  \centering
  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
  \caption{\figtabsize Individual allocation of apartments to students in
    {\residenceblock} that meets the requirements stated before.}
  \label{fig:streetmappathpeople}
\end{figure}

@@

The interval assignment problem to a set system is equivalent to the
consecutive ones problem (COP) in binary matrices\cite{wlh02,
  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
binary matrix in such a way that every column (row) has its {\un}s occur
consecutively. If this is possible the matrix is said to have the COP.
Consecutive ones property (COP) of binary matrices is a well
researched combinatorial problem and has several positive results on
tests for it and computing the COP permutation (i.e. the course
schedule in the above illustration) which will be surveyed later in
this document. Hence we are interested in extensions of COP, more
specifically, the extension of interval assignment problem to tree
path assignment problem (which is illustrated by the study group
accommodation problem).

@@

\section{Background}
\label{sec:background}
\subsection{Matrices with COP}
\label{sec:copmatrices}
As seen earlier, the interval assignment problem (illustrated as the
course scheduling problem in Section~\ref{sec:problem}), is a special
case of the problem we address in this thesis, namely the tree path
labeling problem (illustrated as the study group accommodation
problem). Interval assignment and COP are equivalent
problems. In this section we will see some of the results that
exists in the literature today towards solving the COP problem and
optimization problems surrounding it.

% The interval assignment problem to a set system is equivalent to the
% consecutive ones problem (COP) in binary matrices\cite{wlh02,
%   nsnrs09}.  The COP problem is to rearrange rows (columns) of a
% binary matrix in such a way that every column (row) has its {\un}s occur
% consecutively. If this is possible the matrix is said to have the COP.
% Consecutive ones property (COP) of binary matrices is a well
% researched combinatorial problem and has several positive results on
% tests for it and computing the COP permutation (i.e. the course
% schedule in the above illustration) which will be surveyed later in
% this document. Hence we are interested in extensions of COP, more
% specifically, the extension of interval assignment problem to tree
% path assignment problem (which is illustrated by the study group
% accommodation problem).

@@

The construction of PQ trees in \cite{bl76} draws on the
close relationship of COP matrices to interval graphs. A PQ tree of a
matrix is one that stores all row (column) permutations of the matrix
that give the COP orders (there could be multiple) of the matrix. This
is constructed using an elaborate linear time procedure and is also
a test for planarity\tnote{check check check. both interval
  graph and planarity in this paper?}. % Since this
% result, several improvements have been made to the procedural
% complexity of the algorithm itself and also derivative data structures
% that are easier to construct.
PQR trees is a generalized data structure based on PQ trees
\cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
build PQR trees. \tnote[GTC]{improv in terms of what?}\cite{wlh02}
describes the simpler algorithm for COT. Hsu also invented PC trees
\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
which is claimed to be much easier to implement. \cite{nsnrs09}
describes a characterization of consecutive ones property solely based
on the cardinality properties of the set representations of the
columns (rows); every column (row) is equivalent to a set that has the
row (column) indices of the rows (columns) that have one entries in
this column (row). This is interesting and relevant, especially to
this thesis because it simplifies COT to a great
degree. \tnote[GTC]{it reduces the solution search space. fill in the blanks.}

@@

\subsection{Optimization problems in COP}
\label{sec:optcop}
So far we have been concerned about matrices that have the consecutive
ones property. However in real life applications, it is rare that data
sets represented by binary matrices have COP, primarily due to the
noisy nature of data available. At the same time, COP is not arbitrary
and is a desirable property in practical data representation
\cite{co98,jkckv04,k77}. In this context, there are several
interesting problems when a matrix does not have COP but is ``close''
to having COP or is allowed to be altered to have COP. These are the
optimization problems related to a matrix which does not have
COP. Some of the significant problems are surveyed in this section.

@@

\section{Motivation}
\label{sec:motive}
As seen in Section~\ref{sec:copmatrices}, \cite{nsnrs09} describes a
characterization of consecutive ones property based on the cardinality
properties of the set representations of the columns. This result is
very relevant to this thesis because aside from it simplifying COT to
a great degree, our generalization problem is motivated by their
results.

@@

\section{Summary of results}
\label{sec:results}

We address the computation of TPL\footnote{Tree Path Labeling} for a
given set system and tree in this research. However optimization
problems have not been explored.

Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
necessary and sufficient condition for it to be a feasible path
labeling.  This necessary and sufficient condition can be tested in
polynomial time.  The most interesting consequence is that in our
constructive procedure, it is sufficient to iteratively check if
three-way intersection cardinalities are preserved.  In other words,
in each iteration, it is sufficient to check if the intersection of
any three sets is of the same cardinality as the intersection of the
corresponding paths.  This generalizes the well studied question of
the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}, in
other words the interval assignment problem or consecutive ones problem.

%\item 
Aside from checking if a given TPL is feasible, we also attempt to
compute the TPL for a given set system (hypergraph) and tree, if one
exists. For the latter problem, we have an algorithm for general trees
which runs in exponential time. We present a polynomial time
algorithm for a special class of trees called $k$-subdivided trees and
when the hypergraph $\cF$ is such that all hyperedges have at most
$k+2$ elements.  In spite of this restricted case we consider, we
believe that our results are of significant interest in understanding
the nature of Graph Isomorphism which is polynomial time solvable in
interval graphs and is hard on path graphs. The connection of TPL to graph
isomorphism will be made later in the document
%\end{enumerate}

@@

\subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
\label{sec:feasible}

In this section we give an algorithmic characterization of a
feasibility of tree path labeling.
% Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
% and path system $\cP$ on the given tree $T$. We
% call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
%   (ICPPL)} if it has the following properties.
Consider a path labeling $(\cF, \cl)$ on the
given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
  Preserving Path Labeling (ICPPL)} if it has the following
properties.

% \begin{enumerate}[{\icpplpr\ }i. ]
% \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
%   % \vspace{\topshrink}
% \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
%   for all distinct $S_1, S_2 \in \cF$
%   % \vspace{\topshrink}
% \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
%   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
% \end{enumerate}

\begin{table}[h]
  \centering
  \begin{tabular}[l]{rrcll}
    Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
    Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
    \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
    Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
    |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
    S_2, S_3 \in \cF$ 
  \end{tabular}
  \caption{ICPPL properties}
  \label{tab:icpplpr}
\end{table}

\noindent
The following lemma is useful in subsequent arguments. 
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}

% \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
%   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
%   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
%   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
%   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
%   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
%   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
%   P_3)|$. Thus lemma is proved. \qed
% \end{proof}

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path
% labeling is feasible, then all intersection cardinalities are
% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL \iff input PL
%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
% the sets map to paths, ``preserving'' the path labeling.  The rest
% of the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.

First, we present Algorithm~\ref{perms} called {\tt filter\_1}
% , and
% prove its correctness.  
This algorithm refines the path labeling by
processing pairs of paths in $\cF^\cl$ that share a leaf until no two
paths in the new path labeling share any leaf.
\tnote{write a note summarizing filter 1}
% \begin{algorithm}[h]
%   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
%   \label{perms}
%   \begin{algorithmic}[\lndisplay]
%     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
%     \STATE $j \assign 1$\\
%     \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
%       $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
%       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
%     \{S_1, S_2\})
%     \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
%     \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
%     \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
%     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

%     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
%     \cl_{j-1}(S_2)$
%     \COMMENT {Carry forward the path labeling for all existing sets other than
%       $S_1$, $S_2$}
%     \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
%     \cl_{j-1}(S_2)$ 
%     \COMMENT {Define path labeling for new sets}
%     \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
%     \cl_{j-1}(S_1)$

%     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
%     \label{ln:3waycheck}
%     \STATE {\bf exit} \label{ln:exit1} \\
%     \ENDIF

>>>>

\def \un {\bf 1}
\def \secfirstparatrim {-5mm}
\def \figtabsize {\footnotesize}

@@

\section{Introduction}
\label{sec:intro}
\vspace{\secfirstparatrim}
\tnote{set context: COP - generalization - TPL - no opt. just 2 or 3
  sentences. use the space that the illustration does.}
% \tnote{testing 1 2 3...}
% \tnote[bogus1]{testing 1 2 3...}
% \tnote[bogus2]{testing 1 2 3...}
% \tnote[GTC]{testing 1 2 3...}
% \tnote[bogus4]{testing 1 2 3...}
% \tnote[TH]{testing 1 2 3...}

\subsection{Illustration of the problem}
\label{sec:problem}
\vspace{\secfirstparatrim}
% \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
% \def \GGG {{\em ``{\bf G}irl with Curious Hair''}} 
% \def \BBB {{\em ``{\bf B}rief Interviews with Hideous Men''}}
% \def \TTT {{\em ``Everything and More: A Compact His{\bf T}ory of Infinity''}} 
% \def \xLLL {{\bf L}}
% \def \xGGG {{\bf G}} 
% \def \xBBB {{\bf B}}
% \def \xTTT {{\bf T}}

@@

@@

% \begin{figure}[t]%[htbp]
%   \centering
%   \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
%   \caption{{\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
%   \label{fig:streetmappath}
% \end{figure}
% \begin{figure}[t] % [h]
%   \centering
%   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
%   \caption{Individual allocation of apartments to students in
%     {\residenceblock} that meets the requirements stated before.}
%   \label{fig:streetmappathpeople}
% \end{figure}
%%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
% \begin{figure}[htbp] %[t]%
%  \centering
%   (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
%   (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
%  \label{fig:streetmap}
%   \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
% \end{figure}
% \begin{figure}[htbp] %[t]%
%   \centering
%   \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
%   \caption{\figtabsize Individual allocation of apartments to students in
%     {\residenceblock} that meets the requirements stated before.}
%   \label{fig:streetmappathpeople}
% \end{figure}

@@

The interval assignment problem to a set system is equivalent to the
consecutive ones property (COP) problem in binary matrices\cite{wlh02,
  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
binary matrix in such a way that every column (row) has its {\un}s
occur consecutively. If this is possible the matrix is said to have
the COP.  COP is a well researched combinatorial problem and has
several positive results on tests for it and computing the COP
permutation (i.e. the course schedule in the above illustration) which
will be surveyed later in this document. Hence we are interested in
extensions of COP, more specifically, the extension of interval
assignment problem to tree path assignment problem (which is
illustrated by the study group accommodation problem).

@@

\section{Consecutive Ones Property Testing - a Survey} %{Background}
\label{sec:background}
\vspace{\secfirstparatrim}
In this section, a brief survey of the consecutive ones problem and its
optimization problems is presented.

\subsection{Matrices with COP}
\label{sec:copmatrices}
\vspace{\secfirstparatrim} 
As seen earlier, the interval assignment problem (illustrated as the
course scheduling problem in Section~\ref{sec:problem}), is a special
case of the problem we address in this thesis, namely the tree path
labeling problem (illustrated as the study group accommodation
problem). The interval assignment problem and COP problem are equivalent
problems. In this section we will see some of the results that
exists in the literature today towards solving the COP problem and
optimization problems surrounding it.

@@

The construction of PQ trees in \cite{bl76} draws on the
close relationship of matrices with COP to interval graphs. A PQ tree of a
matrix is one that stores all row (column) permutations of the matrix
that give the COP orders (there could be multiple orders of rows or columns) of the matrix. This
is constructed using an elaborate linear time procedure and is also
a test for planarity\tnote{check check check. both interval
  graph and planarity in this paper?}. % Since this
% result, several improvements have been made to the procedural
% complexity of the algorithm itself and also derivative data structures
% that are easier to construct.
PQR trees is a generalized data structure based on PQ trees
\cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
build PQR trees. \tnote[GTC]{improv in terms of what?}\cite{wlh02}
describes the simpler algorithm for COT. Hsu also invented PC trees
\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
which is claimed to be much easier to implement. \cite{nsnrs09}
describes a characterization of consecutive ones property solely based
on the cardinality properties of the set representations of the
columns (rows); every column (row) is equivalent to a set that has the
row (column) indices of the rows (columns) that have one entries in
this column (row). This is interesting and relevant, especially to
this thesis because it simplifies COT to a great
degree. \tnote[GTC]{it reduces the solution search space. fill in the blanks.}

@@

\subsection{Optimization problems in COP}
\label{sec:optcop}
\vspace{\secfirstparatrim}  
So far we have been concerned about matrices that have the consecutive
ones property. However in real life applications, it is rare that data
sets represented by binary matrices have COP, primarily due to the
noisy nature of data available. At the same time, COP is not arbitrary
and is a desirable property in practical data representation
\cite{co98,jkckv04,k77}. In this context, there are several
interesting problems when a matrix does not have COP but is ``close''
to having COP or is allowed to be altered to have COP. These are the
optimization problems related to a matrix which does not have
COP. Some of the significant problems are surveyed in this section.

@@

\section{Generalization of COP - The Motivation}  %{Motivation}
\label{sec:motive}
\vspace{\secfirstparatrim}  
As seen in Section~\ref{sec:copmatrices}, \cite{nsnrs09} describes a
characterization of consecutive ones property based on the cardinality
properties of the set representations of the columns. This result is
very relevant to this thesis because aside from it simplifying COT to
a great degree, our generalization problem is motivated by their
results.

@@

\def \probdefwidth {0.85\linewidth}
\def \FTPL {{\sc Feasbile Tree Path Labeling}}
\def \CFTPL {\sc Compute Feasible Path Labeling}

\section{Summary of results}
\label{sec:results}
\vspace{\secfirstparatrim}
% We address the computation of TPL\footnote{Tree Path Labeling} for a
% given hypergraph\footnote{The term {\em hypergraph} is used instead
%   {\em set systems} are used interchangably in this
%   section. Hypergraphs that have TPLs on some target tree are called
%   {\em path hypergraphs}.} and target tree in this research. However
% optimization problems in TPL, akin to problems in
% Section~\ref{sec:optcop} for COP, is outside the scope of this thesis.
As we saw earlier by the result in \cite{nsnrs09}, pairwise
intersection cardinality preservation is necessary and sufficient for
an interval assignment to be feasible for a given
hypergraph\footnote{A {\em hypergraph} is an alternate representation
  of a set system and will be used through out this section.} and thus
is a characterization for COP. In our work we extend this
characterization and find that trio-wise intersection cardinality
preservation makes a tree path labeling (TPL) feasible, which is a
generalization of the COP problem. This problem, {\FTPL}\footnote{A
  {\em tree path labeling} $\cl$ is a bijection of paths from the
  target tree $T$ to the hyperedges in given hypergraph $\cF$.}, is
defined as follows.\\

\begin{minipage}[h]{5in}
  % \begin{singlespace}
  {\large \FTPL}\\
  % \hline
  \begin{tabular}[t]{l|l}
    {\tt Input} & 
    \begin{minipage}[t]{\probdefwidth}
      A hypergraph $\cF$ with vertex set $U$, a tree $T$ and a
      bijection $\cl$~$:$~$\cF \rightarrow \cP$, where $\cP$ is a set
      of paths from $T$.\\
    \end{minipage}\\
    {\tt Question} &
    \begin{minipage}[t]{\probdefwidth}
      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
      such that $\phi$ when applied on any hyperedge in $\cF$ will
      give
      the path mapped to it by the given tree path labeling $\cl$.\\
      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
        $S \in \cF$.}
    \end{minipage}\\
  \end{tabular}
  % \end{singlespace}
\end{minipage}\\

We give a necessary and sufficient condition for {\FTPL} to output in
affirmative. This characterization can be checked in polynomial time.
The most interesting consequence is that in our constructive
procedure, it is sufficient to iteratively check if three-way
intersection cardinalities are preserved.  In other words, in each
iteration, it is sufficient to check if the intersection of any three
hyperedges is of the same cardinality as the intersection of the
corresponding paths.  Thus this generalizes the well studied question
of the feasible interval assignment problem which is the special case
when the target tree $T$ is simply a path \cite{wlh02,nsnrs09}.

Aside from checking if a given TPL is feasible, we also solve the
problem of computing a feasible TPL for a given hypergraph and target
tree, if one exists. This problem, {\CFTPL}, is defined as follows. \\

\begin{minipage}[h]{5in}
  % \begin{singlespace}
  {\large \CFTPL}\\
  % \hline
  \begin{tabular}[t]{l|l}
    {\tt Input} & 
    \begin{minipage}[t]{\probdefwidth}
      A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
    \end{minipage}\\

    {\tt Question} &
    \begin{minipage}[t]{\probdefwidth}
      Does there exist a set of paths $\cP$ from $T$ and a bijection
      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
        true} on $(\cF, T, \cl)$.
    \end{minipage}\\
  \end{tabular}
  % \end{singlespace}
\end{minipage}\\

\def \CFTPLKTREE {\sc Compute $k$-subdivided Star Path Labeling}

We present a polynomial time algorithm for {\CFTPL} when the target
tree $T$ belongs to a special class of trees called {\em
  $k$-subdivided stars} and when the hyperedges in the hypergraph
$\cF$ have at most $k+2$ vertices. The examples of $k$-subdivided
stars given in Figure~\ref{fig:ksubstar} makes the definition clear.

%This problem, {\CFTPLKTREE}, is defined as follows. \\
\begin{figure}[htbp]
  \centering
  (a) \includegraphics[scale=0.4]{../img/star.pdf}
  (b) \includegraphics[scale=0.4]{../img/kstar.pdf}
  \caption{\figtabsize Examples of $k$-subdivided stars. (a) $k = 0$ (b) $k = 2$ }
  \label{fig:ksubstar}
\end{figure}

\begin{minipage}[h]{5in}
  % \begin{singlespace}
  {\large \CFTPLKTREE}\\
  % \hline
  \begin{tabular}[t]{l|l}
    {\tt Input} & 
    \begin{minipage}[t]{\probdefwidth}
      A hypergraph $\cF$ with vertex set $U$ such that every hyperedge
      $S \in \cF$ is of cardinality atmost $k+2$ and a $k$-subdivided
      star $T$.\\
    \end{minipage}\\

    {\tt Question} &
    \begin{minipage}[t]{\probdefwidth}
      Does there exist a set of paths $\cP$ from $T$ and a bijection
      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
        true} on $(\cF, T, \cl)$.
    \end{minipage}\\
  \end{tabular}
  % \end{singlespace}
\end{minipage}\\

In spite of this
restricted case we consider, we believe that our results are of
significant interest in understanding the nature of {\sc Graph
  Isomorphism} which is polynomial time solvable in interval graphs
while being hard on path graphs\cite{kklv10}.

\tnote{give gyan on why ktree is so.}

We also have an algorithm for solving {\CFTPL} with no restrictions
which runs in exponential time and explores a partial order theory of
matrices related to {FTPL}. This po theory gives interesting insight
into the problem of finding path labeling from $T$ by decomposing it
into smaller subproblems of finding path labeling of subsets
of $\cF$ from subtrees of $T$.\tnote[TH]{The connection
of TPL to graph isomorphism will be made later in the document}
%\end{enumerate}

@@

% \subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
% \label{sec:feasible} 
% \vspace{\secfirstparatrim}

% In this section we give an algorithmic characterization of a
% feasibility of tree path labeling.
% % Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
% % and path system $\cP$ on the given tree $T$. We
% % call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
% %   (ICPPL)} if it has the following properties.
% Consider a path labeling $(\cF, \cl)$ on the
% given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
%   Preserving Path Labeling (ICPPL)} if it has the following
% properties.

% % \begin{enumerate}[{\icpplpr\ }i. ]
% % \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
% %   % \vspace{\topshrink}
% % \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
% %   for all distinct $S_1, S_2 \in \cF$
% %   % \vspace{\topshrink}
% % \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
% %   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
% % \end{enumerate}

% \begin{table}[h]
%   \centering
%   \begin{tabular}[l]{rrcll}
%     Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
%     Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
%     \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
%     Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
%     |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
%     S_2, S_3 \in \cF$ 
%   \end{tabular}
%   \caption{ICPPL properties}
%   \label{tab:icpplpr}
% \end{table}

% \noindent
% The following lemma is useful in subsequent arguments. 
% \begin{lemma}
%   \label{lem:setminuscard}
%   If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
%   (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
%   \cl(S_3))|$.
% \end{lemma}

% % \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
% %   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
% %   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
% %   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
% %   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
% %   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
% %   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
% %   P_3)|$. Thus lemma is proved. \qed
% % \end{proof}

% In the remaining part of this section we show that $(\cF, \cl)$ is
% feasible if and only if it is an ICPPL and
% Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
% function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
% levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
% the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
% % One direction of this claim isclear: that if a path
% % labeling is feasible, then all intersection cardinalities are
% % preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
% % \annote{has no premature exit condition hence any input will go
% %   through it}{Prove that the filtered sets has ICPPL \iff input PL
% %   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
% % line~\ref{xempty}. It can be easily verified that $X$ cannot be
% % empty if $\cl$ is a feasible path labeling. The reason is that a
% % feasible path labeling has an associated bijection between
% % $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
% % the sets map to paths, ``preserving'' the path labeling.  The rest
% % of the section is devoted to constructively proving that it is
% % sufficient for a path labeling to be an ICPPL and pass the two
% % filtering algorithms.  To describe in brief, the constructive
% % approaches refine an ICPPL iteratively, such that at the end of each
% % iteration we have a ``filtered'' path labeling, and finally we have
% % a path labeling that defines a family of bijections from $supp(\cF)$
% % to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.

% First, we present Algorithm~\ref{perms} called {\tt filter\_1}
% % , and
% % prove its correctness.  
% This algorithm refines the path labeling by
% processing pairs of paths in $\cF^\cl$ that share a leaf until no two
% paths in the new path labeling share any leaf.
% \tnote{write a note summarizing filter 1}
% % \begin{algorithm}[h]
% %   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
% %   \label{perms}
% %   \begin{algorithmic}[\lndisplay]
% %     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
% %     \STATE $j \assign 1$\\
% %     \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
% %       $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
% %       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
% %     \{S_1, S_2\})
% %     \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
% %     \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
% %     \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
% %     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

% %     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
% %     \cl_{j-1}(S_2)$
% %     \COMMENT {Carry forward the path labeling for all existing sets other than
% %       $S_1$, $S_2$}
% %     \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
% %     \cl_{j-1}(S_2)$ 
% %     \COMMENT {Define path labeling for new sets}
% %     \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
% %     \cl_{j-1}(S_1)$

% %     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
% %     \label{ln:3waycheck}
% %     \STATE {\bf exit} \label{ln:exit1} \\
% %     \ENDIF
